Adaptive broadcast tree for multicores

### 2013-04-30 paper.tex
After meeting Kornilios

% ==================================================
\setcounter{section}{0} % Start sections with 1, not 0
\begin{document}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related work}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The multicore network model}

@@

\subsection{Node characteristics}

@@

\begin{description}
\item[synchrony] Parts of a multicore machine are synchronous. Hence,
  one characteristic of a node is an identifier for the synchrony
  cluster of a node.
\item[$t_{send}$] Time required to send a message, i.e.\ marshalling
  and queuing the message
\item[$t_{recv}$] Time required to receive a message, i.e.\
  dequeuing, unmarshalling and demultiplexing a channel.
\item[failure] Can nodes fail? If so, how? (Byzantine, \ldots)
\end{description}

\subsection{Link characteristics}

@@

\begin{description}
\item[latency] (quantitative) between each pair of nodes (per-edge)
\item[breakdown] (quantitative) of message send cost:
  dominating propagation time vs.\ dominating send and receive time on
  multicores (per-node)
\item[bounded delivery]
\item[reliability] (loss, link failure, in-order delivery, phantoms,
  corruption, duplicated messages): 
  multicore interconnect networks often are
%\item[Security] \& and failure detection (and the cost for that)
\end{description}

@@

\subsection{Acquiring the model}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Example: Atomic broadcast
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Atomic Broadcast}

We build our atomic broadcast based on an overlay network connecting
all nodes of a given multicore machine.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Tree}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Divide and conquer}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Clustering}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Minimum-Spanning Tree}

@@

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Experimental results}

\subsection{Simulation in model}

@@

\subsection{Real hardware}

\paragraph{Current numbers} Current numbers are listed in
Table~\ref{tab:bc_measurements}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Future work}

\begin{itemize}
\item Arguing that group communication/atomic broadcast and showing
  that we can do it well might be enough (appeal to atomic broadcasts
  as the foundation for other distributed algorithms, and the MPI
  primitives, that can/need to be mapped to group-communication (have
  a list of them) 
\item Need higher-level applications, such as:
  \begin{itemize}
  \item capability system? \stefan{Ask Simon how expensive revocation
      currently is}
  \item a database with replication and consistency maintenance
    \begin{itemize}
    \item SharedDB? (buffers between operators, but no synchronization
      except for access to buffers, which are multiple writers, one
      reader)
    \item Crescando (some kind of state machine replication)
    \end{itemize}
  \end{itemize}
\item Machine model for topology aware tree
  \begin{itemize}
  \item Talked to Pravin about this a bit
  \end{itemize}
\item MPI collectives (as in~\cite{Tu2008})
\end{itemize}

\subsection{Octopus}

One idea is to make Octopus a decentralized application. Octopus is
a coordination service implemented for the Barrelfish Multikernel
operating system. It is, however, centralized.

The idea is to extend Octopus such that it can be replicated within a
multicore machine. This will retain scalability for future manycore
machines and potentially outperform a centralized version of Octopus
especially in contented systems.

@@

The question is whether a key-value store as a backend for
synchronization primitives even makes sense.

@@

\subsubsection{Load balancing}

>>>>

\setcounter{section}{0} % Start sections with 1, not 0
\begin{document}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related work}

@@

Alistarh\etal~\cite{Alistarh2012} show composition of shared-memory
algorithms that perform efficiently under different conditions at
negligible overhead. A similar approach can potentially be applied to
deal with diversity of multicore machines. However, we would have
to extend it to support both shared-memory and message-passing
implementations.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The multicore network model}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Node characteristics} 
\label{sec:model_nodes}

@@

\begin{description}
\item[synchrony] Parts of a multicore machine are synchronous. Hence,
  one characteristic of a node is an identifier for the synchrony
  cluster of a node.
\item[$t_{send}$] Time required to send a message, i.e.\ marshalling
  and queuing the message
\item[$t_{recv}$] Time required to receive a message, i.e.\
  dequeuing, unmarshalling and demultiplexing a channel.
\item[failure] Can nodes fail? If so, how? (Byzantine, crash)
\end{description}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Link characteristics} 
\label{sec:model_links}

@@

\begin{description}
\item[latency] (quantitative) between each pair of nodes (per-edge)
\item[breakdown] (quantitative) of message send cost:
  dominating propagation time vs.\ dominating send and receive time on
  multicores (per-node)
\item[bounded delivery]
\item[reliability] (loss, link failure, in-order delivery, phantoms,
  corruption, duplicated messages): 
  multicore interconnect networks often are
\end{description}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Hierarchical model}
\label{sec:model_hierarchy}

Multicore machines are hierarchical. Processor cores are grouped in
NUMA nodes with shared caches. These NUMA nodes are then connected to
each by increasingly complex interconnect networks.

We encode NUMA node affinity in our model. This allows to break down
programming of a multicore machine into two distinct instances. We
solve one instance of the problem between NUMA nodes, and then, in a
second step, within each NUMA node. These two steps can be executed
independently and different algorithms can be used in either case.

If cores on the same NUMA node share a cache, it makes sense to
leverage this hardware feature to implement shared-memory algorithms
on top of it. %
Communication across NUMA nodes is realized sending messages across
the packet-based interconnect network. Using message-passing based
implementations is the natural choice. %
Shared memory algorithms and message-passing based algorithms can be
composed together to program hierarchical machines similarly to what
was previously shown for combinations of shared-memory
algorithms~\cite{Alistarh2012}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Acquiring the model}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Atomic Broadcast}

We build our atomic broadcast based on an overlay network connecting
all nodes of a given multicore machine. First, we discuss what
kind of overlay networks make sense on multicores.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Tree overlay}

@@

%-------------------------------------------------
\subsubsection{Divide and conquer}

@@

%-------------------------------------------------
\subsubsection{Clustering}

@@

%-------------------------------------------------
\subsubsection{Minimum-Spanning Tree}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Rings}

\stefan{Just a random collection of ideas}

%-------------------------------------------------
\subsubsection{Load balancing in a tree}

@@

\stefan{This section feels out of place. Rewrite it and put it
  somewhere else}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Evaluation}

\stefan{Compare to pure shared-memory implementations, otherwise
  reviewers might argue that there is no need for distributed
  algorithms.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Atomic broadcast}

%-------------------------------------------------
\subsubsection{Simulation}

@@

%-------------------------------------------------
\subsection{Real hardware}

Current numbers are listed in Table~\ref{tab:bc_measurements}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Future work}

\begin{itemize}
\item Arguing that group communication/atomic broadcast is important
  and showing that we can do it well might be enough (appeal to atomic
  broadcasts as the foundation for other distributed algorithms, and
  the MPI primitives, that can/need to be mapped to
  group-communication (have a list of them)
\item Need higher-level applications, such as:
  \begin{itemize}
  \item capability system? \stefan{Ask Simon how expensive revocation
      currently is}
  \item a database with replication and consistency maintenance
    \begin{itemize}
    \item SharedDB? (buffers between operators, but no synchronization
      except for access to buffers, which are multiple writers, one
      reader)
    \item Crescando (some kind of state machine replication)
    \end{itemize}
  \end{itemize}
\item MPI collectives (as in~\cite{Tu2008})
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Octopus}

Octopus is a coordination service implementation for the Barrelfish
Multikernel operating system. It is, however, centralized. The idea is
to make Octopus a decentralized application.

For that, we extend Octopus such that it can be replicated within a
multicore machine. This will retain scalability for future manycore
machines and potentially outperform a centralized version of Octopus
especially in contented systems.

@@

%--------------------------------------------------
\subsubsection{Key-value store}

The question is whether a key-value store as a backend for
synchronization primitives even makes sense. As we want to show
efficiency, we probably rather want to implement new synchronization
algorithms from scratch.

@@
